Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems

### 2024-06-26 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA part 2

\begin{equation}
\label{eq:01}
  \Phi(y) = (f_1 (y),f_2 (y), \dots, f(y)) \to \min, y \in D
\end{equation}

\begin{equation}
\label{eq:02}
    D=\{y \in R^n : a_i \leq y_i \leq b_i, 1 \leq i \leq N\}
\end{equation}

\begin{equation}
\label{eq:03}
|f_i (y') - f_i (y'')| \leq L_i \|y' - y''\| ,y',y'' \in D, 1 \leq i \leq s,
\end{equation}

>>>>

The multicriterial optimization (MCO) problem can be formulated as the minimization problem of the vector-function. Let us introduce the vector-function (vector criterion)

\begin{equation}
\label{eq:01}
  \Phi(y) = (f_1 (y),f_2 (y), \dots, f_s(y))
\end{equation}
where scalar functions $f_i(y)$, $1 \leq i \leq s$, are considered as the partial efficiency criteria, $y=(y_1,y_2, \dots ,y_N)$ is the vector of independent parameters belonging to the feasible domain $D \in R^N$ that is the hyperparallelepiped
\begin{equation}
\label{eq:02}
    D=\{y \in R^N : a_i \leq y_i \leq b_i, 1 \leq i \leq N\}
\end{equation}
in the $N$-dimensional Euclidean space, where $a_i$, $b_i$, $1 \leq i \leq N$, are constant values.

The best value of the partial criterion $f_i(y)$, $1 \leq i \leq s$, is supposed to correspond to the global minimum of this criterion in the domain $D$.

We will denote this problem in the form
\begin{equation}
\label{eq:02a}
  \Phi(y) = (f_1 (y),f_2 (y), \dots, f_s(y)) \to \min, y \in D
\end{equation}

In the framework of this study, one of the most complicated classes of the above statement is considered in which the criteria $f_i(y)$, $1 \leq i \leq s$, are multiextremal and presented as the ``black-box'' functions. Moreover, these criteria are supposed to satisfy the Lipschitz condition

\begin{equation}
\label{eq:03}
|f_i (y') - f_i (y'')| \leq L_i \|y' - y''\| ,y',y'' \in D, 1 \leq i \leq s,
\end{equation}
where Lipschitz constants $L_i>0$, $1 \leq i \leq s$, are unknown \textit{a priopi}.

Without loss of generality, the values of partial criteria will be supposed to be positive. If a partial criterion does not satisfy this requirement, it is easy to build a new criterion by means of adding to it a sufficiently large positive constant so that replacing in the problem (\ref{eq:02a}) the initial criterion with the modified one will not change its solution. Note that such constant can be selected in the course of optimization \cite{ML_MCO_2023}.

As it was mentioned above, the partial criteria contradictoriness leads to the necessity to introduce the set of non-dominated points   called Pareto set as the general solution of the multicriterial problem \cite{Miettinen1999,Ehrgott2005,Pardalos2017}. Hereinafter, any point of the Pareto set will be called a partial solution or just an effective point.
